Figure two

### 2025-10-27 draftFiguresPaper.Rmd
cleanUp data code Fig2

# 2) get total surveys by country
surveys_by_country <- surveyCountries_long_clean %>%
  group_by(iso3, .drop = F) %>% 
  count(iso3, name = "n_surveys") %>%
  arrange(desc(n_surveys))

@@

>>>>

# 2) get total surveys by country
surveys_by_country <- surveyCountries_long_clean %>%
  filter(Inclusion == 1) %>% 
  group_by(iso3, .drop = F) %>% 
  count(iso3, name = "n_surveys") %>%
  arrange(desc(n_surveys))

@@

```{r long classification dataset - paper}

library(tidyverse)
library(countrycode)

datasetCountries <- datasetCountries %>% 
  dplyr::select(ID, countryPartialF) %>% 
  rename(country = countryPartialF)

datasetClassificationCountries <- merge(datasetClassification, datasetCountries,
                                        by = "ID") %>% 
  dplyr::select(-reader)

datasetClassificationCountries_long$countries <- ifelse(datasetClassificationCountries_long$countries == "Turkyie", "Turkey",
                                                        ifelse(datasetClassificationCountries_long$countries %in% c("Wales", "England"), "United Kingdom",
                                                               datasetClassificationCountries_long$countries))
# fix country names 
# 1. Direct mapping
datasetClassificationCountries_long_clean <- datasetClassificationCountries_long %>%
  mutate(
    iso3 = countrycode(countries,
                       origin      = "country.name",
                       destination = "iso3c")
  ) %>% 
  filter(countries %ni% c("", " ", "Kosovo", "Southeast Asia"))

# add the year
datasetYear <- read.csv("dataNew/datasetToClassifyGPT.csv") %>% 
  dplyr::select(ID, Year)

df_cat_long <- datasetClassificationCountries_long_clean %>%
  pivot_longer(
    cols = c(Perceptions, Priority, Policy, Health, Behavior),
    names_to  = "category",
    values_to = "flag"
  ) %>%
  # keep only “paper belongs to this category”
  filter(flag == 1) %>%
  dplyr::select(-flag, -None)

write.csv(df_cat_long, "output/paperClassifYear.csv", row.names = F)

```{r long classification dataset - survey}

library(tidyverse)
library(countrycode)

df_cat_long <- surveyCountries_long_clean %>%
  pivot_longer(
    cols = c(Perceptions, Priority, Policy, Health, Behavior),
    names_to  = "category",
    values_to = "flag"
  ) %>%
  # keep only “paper belongs to this category”
  filter(flag == 1) %>%
  dplyr::select(-flag)

write.csv(df_cat_long, "output/surveyClassifYear.csv", row.names = F)

### 2025-10-27 draftFiguretwo_Ella.Rmd
cleanUp data code Fig2

#-------------------------------------------------------------------------------
# Load the data
#-------------------------------------------------------------------------------
papers <- read.csv2("dataNew/data_fig2_papers.csv", 
                    header = T, sep = ",", dec = ".")

surveys <- read.csv2("dataNew/data_fig2_surveys.csv", 
                     header = T, sep = ",", dec = ".")

@@

## Clean the data

@@

@@

# reshape to long format and count topics
papers_long <- papers %>%
  pivot_longer(
    cols = c(Perceptions, Behavior, Priority, Policy, Health, None),
    names_to = "Topic",
    values_to = "Value"
  ) %>%
  filter(Value == 1) %>%
  group_by(Topic) %>%
  summarise(n = n()) %>%
  mutate(share = n / sum(n))

@@

## Different visualisation
install.packages("treemapify")
library(treemapify)

@@

# reshape to long format and count topics
surveys_long <- surveys %>%
  pivot_longer(
    cols = c(Perceptions, Behavior, Priority, Policy, Health),
    names_to = "Topic",
    values_to = "Value"
  ) %>%
  filter(Value == 1) %>%
  group_by(Topic) %>%
  summarise(n = n()) %>%
  mutate(share = n / sum(n))

@@

install.packages("ggnewscale")
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggnewscale)

# 1. Prepare topic-level data (shares within income group)
topics_data <- papers %>%
  pivot_longer(
    cols = c(Perceptions, Behavior, Priority, Policy, Health, None),
    names_to = "Topic",
    values_to = "Value") %>%
  filter(Value == 1, !is.na(Income.group)) %>%
  group_by(Topic, Income.group) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(Income.group) %>%
  mutate(share = n / sum(n)) %>%
  ungroup() %>%
  mutate(Income.group = factor(
    Income.group,
    levels = c("Low income", "Lower middle income", "Upper middle income", "High income")))

# 2. Prepare totals row data (absolute counts)
totals_data <- papers %>%
  filter(!is.na(Income.group)) %>%
  count(Income.group, name = "total")

@@

# 1. Prepare topic-level data (shares within income group)
topics_data_s <- surveys %>%
  pivot_longer(
    cols = c(Perceptions, Behavior, Priority, Policy, Health),
    names_to = "Topic",
    values_to = "Value") %>%
  filter(Value == 1, !is.na(Income.group)) %>%
  group_by(Topic, Income.group) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(Income.group) %>%
  mutate(share = n / sum(n)) %>%
  ungroup() %>%
  mutate(Income.group = factor(
    Income.group,
    levels = c("Low income", "Lower middle income", "Upper middle income", "High income")))

# 2. Prepare totals row data (absolute counts)
totals_data_s <- surveys %>%
  filter(!is.na(Income.group)) %>%
  count(Income.group, name = "total_items")

@@

#-------------------------------------------------------------------------------
# get exposure groups for papers and surveys
#-------------------------------------------------------------------------------
papers$categoryPollution <- cut(papers$pm25Avg, breaks = c(0, 10, 15, 25, 35, 100), 
                                labels = c("Very low", "Low", "Medium", "High", "Very high"), 
                                right = FALSE)

surveys$categoryPollution <- cut(surveys$pm25Avg, breaks = c(0, 10, 15, 25, 35, 100), 
                                labels = c("Very low", "Low", "Medium", "High", "Very high"), 
                                right = FALSE)

@@

# 1. Prepare topic-level data (shares within exposure group)
topics_data_poll <- papers %>%
  pivot_longer(
    cols = c(Perceptions, Behavior, Priority, Policy, Health, None),
    names_to = "Topic",
    values_to = "Value") %>%
  filter(Value == 1, !is.na(categoryPollution)) %>%
  group_by(Topic, categoryPollution) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(categoryPollution) %>%
  mutate(share = n / sum(n)) %>%
  ungroup() %>%
  mutate(categoryPollution = factor(
    categoryPollution,
    levels = c("Very low", "Low", "Medium", "High", "Very high")))

# 2. Prepare totals row data (absolute counts)
totals_data_poll <- papers %>%
  filter(!is.na(categoryPollution)) %>%
  count(categoryPollution, name = "total")

@@

# 1. Prepare topic-level data (shares within exposure group)
topics_data_poll_s <- surveys %>%
  pivot_longer(
    cols = c(Perceptions, Behavior, Priority, Policy, Health),
    names_to = "Topic",
    values_to = "Value") %>%
  filter(Value == 1, !is.na(categoryPollution)) %>%
  group_by(Topic, categoryPollution) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(categoryPollution) %>%
  mutate(share = n / sum(n)) %>%
  ungroup() %>%
  mutate(categoryPollution = factor(
    categoryPollution,
    levels = c("Very low", "Low", "Medium", "High", "Very high")))

# 2. Prepare totals row data (absolute counts)
totals_data_poll_s <- surveys %>%
  filter(!is.na(categoryPollution)) %>%
  count(categoryPollution, name = "total_items")

>>>>

#-------------------------------------------------------------------------------
# Load the data
#-------------------------------------------------------------------------------
papers <- read.csv2("output/paperClassifYear.csv", 
                    header = T, sep = ",", dec = ".")

surveys <- read.csv2("output/surveyClassifYear.csv", 
                     header = T, sep = ",", dec = ".")

dataSummary <- read.csv("output/datasetPlots.csv")

dataSummary$categoryPollution <- ifelse(
  !is.na(dataSummary$categoryPollution),
  paste0(toupper(substr(dataSummary$categoryPollution, 1, 1)),
         substr(dataSummary$categoryPollution, 2, nchar(dataSummary$categoryPollution))),
  NA
)

dataSummary$categoryPollution <- factor(
  dataSummary$categoryPollution,
  levels = rev(c(
    "Very high",
    "High",
    "Medium",
    "Low",
    "Very low"
  )),
  ordered = TRUE
)

papers <- merge(papers, dataSummary, by = "iso3", all.x = T, all.y = F)
surveys <- merge(surveys, dataSummary, by = "iso3", all.x = T, all.y = F)

@@

## Clean the data (would skip this?)

@@

# install.packages("treemapify")
library(tidyverse)
library(scales)
library(treemapify)

@@

# count category
papers_long <- papers %>%
  count(category, name = "n") %>%   # counts per category
  mutate(share = n / sum(n)) %>%    # sum(n) is the grand total here
  rename(Topic = category)

@@

## Different visualisation

@@

surveys_long <- surveys %>%
  count(category, name = "n") %>%   # counts per category
  mutate(share = n / sum(n)) %>%    # sum(n) is the grand total here
  rename(Topic = category)

@@

# install.packages("ggnewscale")
library(tidyverse)
library(scales)
library(ggnewscale)

# 1. Prepare topic-level data (shares within income group)
topics_data <- papers %>%
  count(Income.group, category, name = "n") %>%   # counts per category
  rename(Topic = category) %>% 
  filter(!is.na(Income.group)) %>% 
  mutate(Income.group = factor(
    Income.group,
    levels = c("Low income", "Lower middle income", "Upper middle income", "High income"))) %>% 
  group_by(Income.group) %>% 
  mutate(share = n / sum(n))   # sum(n) is the grand total here

# 2. Prepare totals row data (absolute counts)
totals_data <- dataSummary %>%
  filter(!is.na(Income.group)) %>%
  group_by(Income.group) %>% 
  summarize(total = sum(n_papers, na.rm=T))

@@

# 1. Prepare topic-level data (shares within income group)
topics_data_s <- surveys %>%
  count(Income.group, category, name = "n") %>%   # counts per category
  rename(Topic = category) %>% 
  filter(!is.na(Income.group)) %>% 
  mutate(Income.group = factor(
    Income.group,
    levels = c("Low income", "Lower middle income", "Upper middle income", "High income"))) %>% 
  group_by(Income.group) %>% 
  mutate(share = n / sum(n))   # sum(n) is the grand total here

# 2. Prepare totals row data (absolute counts)
totals_data_s <- dataSummary %>%
  filter(!is.na(Income.group)) %>%
  group_by(Income.group) %>% 
  summarize(total_items = sum(n_surveys, na.rm=T))

@@

#-------------------------------------------------------------------------------
# get exposure groups for papers and surveys
#-------------------------------------------------------------------------------

## now is already in the papers dataset, could change values based on pollution value (pm25$categoryPollution <- cut(papers$pm25Avg, breaks = c(0, 10, 15, 25, 35, 100), 
#                                 labels = c("Very low", "Low", "Medium", "High", "Very high"), 
#                                 right = FALSE)

# surveys$categoryPollution <- cut(surveys$pm25Avg, breaks = c(0, 10, 15, 25, 35, 100), 
#                                 labels = c("Very low", "Low", "Medium", "High", "Very high"), 
#                                 right = FALSE)

@@

# 1. Prepare topic-level data (shares within exposure group)
topics_data_poll <- papers %>%
  count(categoryPollution, category, name = "n") %>%   # counts per category
  rename(Topic = category) %>% 
  filter(!is.na(categoryPollution)) %>% 
  group_by(categoryPollution) %>% 
  mutate(share = n / sum(n))   # sum(n) is the grand total here

# 2. Prepare totals row data (absolute counts)
totals_data_poll <- dataSummary %>%
  filter(!is.na(categoryPollution)) %>%
  group_by(categoryPollution) %>% 
  summarize(total = sum(n_papers, na.rm=T))

# 3. Define consistent factor levels for the y-axis
topic_levels <- c("Total papers", unique(topics_data_poll$Topic))

@@

# 1. Prepare topic-level data (shares within exposure group)
topics_data_poll_s <- surveys %>%
  count(categoryPollution, category, name = "n") %>%   # counts per category
  rename(Topic = category) %>% 
  filter(!is.na(categoryPollution)) %>% 
  group_by(categoryPollution) %>% 
  mutate(share = n / sum(n))   # sum(n) is the grand total here

# 2. Prepare totals row data (absolute counts)
totals_data_poll_s <- dataSummary %>%
  filter(!is.na(categoryPollution)) %>%
  group_by(categoryPollution) %>% 
  summarize(total_items = sum(n_surveys, na.rm=T))

# 3. Define consistent factor levels for the y-axis
topic_levels_s <- c("Total items", unique(topics_data_poll_s$Topic))
